---
title: "Assignment 6. Web Scraping"
author: "Anabel Berjon Sanchez"
date: "15 Oct 2020"
output: html_document
---


### 0. Preparation: Load packages

```{r, message = FALSE}
library(tidyverse) 
library(rvest)
library(stringr)
```



<br>

### 1. The ultimate regular expression (1pt)

Find a regular expression that matches anything of any length! Prove that it works using several examples.


```{r, eval = TRUE}
# enter your R code here

y<-"Europe also is bracing for a dark winter as the pandemic has entered a dangerous phase across the continent"

y<-c("Europe also is bracing for a dark winter as the pandemic has entered a dangerous phase across the continent", "supercalifragilisticexpialidocious") 

# Match all words with 1,2 and 3 letters
length1<- str_extract_all(y,"\\b[a-zA-Z0-9]{1,3}\\b")
length1

#Match all words with 4 letters
lenght2<-str_extract_all(y,"\\b[a-zA-Z]{4}\\b")
lenght2

#Match all words longer than 5 letters
lenght3<-str_extract_all(y,"\\b[a-zA-Z]{6,}\\b")
lenght3

x<-"It's supercalifragilisticexpialidocious. Even though the sound of it. Is. something quite atrocious. If you say it loud enough. You'll always sound precocious. Supercalifragilisticexpialidocious"

#Match all words longer than 12 letters
leght4<-str_extract_all(x,"\\b[a-zA-Z]{12,}\\b")
leght4
```


<span style="color:blue">
If you want to add a comment outside the R code block, use this <span> block! (just copy the <span>...</span> lines for future comments in the Rmd file.)
</span>

<br>

### 2. Finding strings that are matched by regular expressions (2pts)

Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

a. [0-9]+\\$
b. b[a-z]{1,4}
c. .*?\\.txt$
d. \\d{2}/\\d{2}/\\d{4}
e. <(.+?)>.+?</\\1>

```{r}
# Example
str_extract_all("Phone 150$, TV 690$", "[0-9]+\\$") # Answer: this regex describes prices in dollars

# enter your R code here to solve b. to e.

#B: this regex extracts the words in the string which have the length of 1 to 4 letters. 
stringa<-c("This is President Trump shifted the way both parties talk about trade, immigration and deficits")
stringa<-str_extract_all(stringa, "\\b[a-z]{1,4}")
stringa

#C:Matchs strings that end with ‘.txt’
stringc<-c("Mydoc.exe", "Mydoc2.exe", "Mydoc3.txt")
str_extract_all(stringc,".*?\\.txt$")

#D: Matchs digits with date format of 00/00/0000
stringd<-c("000/3/123","10/02/1993","14/9/1960 ","28/12\1989", "mom" )
stringd <- str_extract(stringd,"\\d{2}/\\d{2}/\\d{4}")
stringd

#E: Matchs any string which starts and ends with tag (<>)
stringe<- c("<tag> anabel/>","<tag>berjon</tag>","<tag sanchez /tag>")
stringe <- str_extract(stringe,"<(.+?)>.+?</\\1>")
stringe
```


<br>

### 3. Manipulating an email address (2pts)

Consider the mail address  datalover89[at]aol[dot]com. Now, complete the following tasks.

a. Transform the string to a standard mail format (i.e., replace the square brackets with the appropriate symbols) using regular expressions.
b. Now extract the local-part ("datalover89") and the domain ("aol") separately. 

```{r}
#A: Replace email address with regex: datalover89[at]aol[dot]com
email <- "datalover89[at]aol[dot]com"%>%
str_replace("\\[at]", "@" )%>%
str_replace("\\[dot]", "\\.")

#B:
domain<-gsub(".*@|\\..*", "", email)
domain
local<-gsub("@.*", "", email)
local
```

<br>

### 4. A secret message (2pts)
The following code hides a secret message. Crack it with R and regular expressions. Once you have cracked it, try to collapse the solution in one single string. <i>Hint: Some of the characters are more revealing than others!</i>

```{r}
secret <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.RqoE65fGEa1otfb7wXm24k.6t3sH9zqe5fy89n6Ed5t9kc4fR905gmc4Ogxo5nhk!gr"

# enter your R code here

secret <- c("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")

message <- str_extract_all(secret,"[[:upper:]]")

message
```



<br>

### 5. Scraping newspaper headlines (3pts)

Use Selectorgadget and R to scrape the article headlines from https://www.theguardian.com/international. 

a. Present the first 6 observations from the uncleaned vector of scraped headlines.

b. Tidy the text data (e.g., remove irrelevant characters if there are any, and get rid of duplicates).

c. Identify the 5 most frequent words in all headlines. (Hint: use a string processing function from the stringr package to split up the headings word by word, and use an empty space, " ", as splitting pattern.)


```{r}
guardian_url <- "https://www.theguardian.com/international"
html <- read_html(guardian_url)
cast <- html_nodes(html, ".js-headline-text , .fc-sublink__link")

# First 6 observations 
cast <- cast[1:6]
cast

# Tidy
cast <- html_text(cast, trim = TRUE)
cast

# 4 most frequent words
cast1 <- str_split(cast, " ")

cast1 <- unlist(cast1) 
cast1 <- sapply(cast1, paste0, collapse=" ")

table(cast1)

sort(table(cast1), decreasing=T)[1:4]
```


<br>

### 6. Towers of the world (3pts)

Scrape the table "Towers proposed or under construction" from https://en.wikipedia.org/wiki/List_of_tallest_towers.

a. Present the first 6 rows of the generated data frame.

b. How many of those buildings are planned for observation purposes?

c. What is the sum of the planned pinnacle height of all those towers? 

```{r}

tow_url <- read_html("https://en.wikipedia.org/wiki/List_of_tallest_towers")
tow_all <- html_table(tow_url, header = TRUE, fill = TRUE)
tow <- tow_all[[7]]

# First 6 observations 
tow[1:6]

#B: Correcting craping error 
tow$Function[tow$Function == "proposed"] <- "observation / telecommunications" 

# Buildings planned for observation purposes 
tow %>% filter(grepl("observation", Function))

# Answer:8 buildings are planned for observation purposes: 3 are for observation and/or telecommunications, and 1 for observation and/orleisure and/or exhibition.

# C: Sum of planned pinnacle height of all towers
tow1 <- tow 
tow1$`Pinnacle height` <- gsub("m", "", 
                              tow1$`Pinnacle height`)
tow1 <- tow1 %>% mutate(`Pinnacle height` =
                as.numeric(tow1$`Pinnacle height`))
sumpin <- sum(tow1$`Pinnacle height`)
sumpin
```

<br>

